---
title: "Bioprint2020 Analysis : all sets combined "
output: 
  html_notebook:
    code_folding: hide
    df_print: paged
---

## All sets combined analysis {.tabset}

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

This notebook describes the analysis comparing all datasets : handmade/manual 'hand', single deposition organoids (R0), and line organoids (R40).

Data have been normalised using SCTransform and integrated.

### Load data

```{r}
library(Seurat)
library(ggplot2)
library(clustree)
library(future)
library(edgeR)
library(dplyr)

source("functions/export_marker_xls.R")
source("functions/sub_cluster.R")
source("functions/get_props_table.R")
source("functions/do_pairwise_edgeR.R")
source("functions/plot_by_set.R")


if (supportsMulticore()){
  
plan('multiprocess', workers=16)
options(future.globals.maxSize = 8000 * 1024^2)
print('Notebook processed on Meerkat HPC using parallel processing')

} else {
  print('multi-core disabled on Rstudio')
}

date()

condition.string <- 'all'

```


```{r load-data}

if (!file.exists("output/integrated-all-clust.RDS")){

int <- readRDS("output/integrated-all.RDS")   
int <- RunPCA(int)
int <- RunUMAP(int, dims=1:30)
int <- FindNeighbors(int, dims = 1:30, verbose = F)
int <- FindClusters(int, resolution = seq(0.1, 2, 0.1), verbose = F)

saveRDS(int, "output/integrated-all-clust.RDS")

} else {
  int <- readRDS("output/integrated-all-clust.RDS")
}


#add metadata fields for condition and replicate

int@meta.data['orig.name'] <- int@meta.data$orig.ident
int@meta.data$orig.ident <- factor(int@meta.data$orig.name, 
            levels = c("hand","dots","line"), 
            labels = c("Man","R0", "R40"))

table(int@meta.data$orig.ident)

int@meta.data['rep'] <- paste(int@meta.data$orig.ident, 
                              substring(int@meta.data$hash.ID, 1,5), sep = "_")


```

```{r plot-cell-numbers-by-set}

sets.plt <- as.data.frame(table(int@meta.data$hash.ID, int@meta.data$orig.ident))

sets.plt$Var1 <- factor(substr(sets.plt$Var1, 5, 5))

sets.plt$Var2 <- factor(sets.plt$Var2, labels = c("Man","R0","R40"))

sets.plt.obj <- ggplot(sets.plt, aes(x = Var2, y = Freq, shape = Var1)) + geom_point(size = 4) +
  xlab(label = "") + ylab(label = "") + theme_classic() + theme(legend.title = element_blank())

ggsave(filename = "plots/all/HTO-numbers-by-set.png", plot = sets.plt.obj, dpi = 300, units = "mm", width = 60, height = 60)

```



Display basic QC info

```{r}
print("cells per dataset: ")
table(int@meta.data$orig.ident)
print("total genes: ")
nrow(int@assays$RNA@counts)
print("Median genes per cell: ")
median(int@meta.data$nFeature_RNA)
print("Median UMI counts per cell: ")
median(int@meta.data$nCount_RNA)
```

```{r CC-scoring}
genes <- Seurat::cc.genes.updated.2019

int <- CellCycleScoring(int, s.features = genes$s.genes, g2m.features = genes$g2m.genes)
```

### Global Clustering

Clustree for multiple resolutions

```{r clustree, fig.height=8, fig.width=10}
clustree(int)

```

Using clustering resolution 0.4

```{r fig.height=4, fig.width=6}

int <- SetIdent(int, value = int@meta.data$integrated_snn_res.0.4)
DefaultAssay(int) <- "RNA"
int <- NormalizeData(int, verbose = F)

DimPlot(int, label = T)

```


```{r cluster-QC-plots}
p1 <- VlnPlot(int, features = "nCount_RNA", pt.size = 0)
p2 <- VlnPlot(int, features = "nFeature_RNA", pt.size = 0)
p3 <- VlnPlot(int, features = "percent.mito", pt.size = 0)
p4 <- VlnPlot(int, features = "scrublet_score", pt.size = 0)

gridExtra::grid.arrange(p1, p2, p3, p4)

```



```{r raw-numbers-and-proportions}

print("raw numbers / proportions")
cbind(table(int@active.ident, int$orig.ident), prop.table(table(int@active.ident, int$orig.ident), margin=2))


```

```{r global-cluster-proportions}

df <- prop.table(table(int@active.ident, int$rep), margin = 2)
df <- df %>% as.data.frame() 

df['condition'] <- sapply(as.character(df$Var2), FUN = function(x){return(strsplit(x, split = "_")[[1]][1])})

df <- df %>% group_by(Var1) %>% mutate(norm.prop = Freq / mean(Freq))

df %>% ggplot(aes(x=condition,y=norm.prop)) + geom_point() + facet_wrap(~Var1) + geom_boxplot(alpha = 0.1)

```



### Global Markers

Get markers for global clusters

```{r global-markers}

if(!file.exists(paste0('output/', condition.string, '-all-cluster-markers.RDS'))){
  markers <- FindAllMarkers(int, logfc.threshold = 0.25, only.pos = TRUE)
  saveRDS(markers, paste0('output/', condition.string, '-all-cluster-markers.RDS'))
  export_marker_xls(markers, xls.path = paste0("output/", condition.string, "-all_cluster_markers.xlsx"))
 
}

markers <- readRDS(paste0('output/', condition.string, '-all-cluster-markers.RDS'))

```

```{r}
markers %>% group_by(cluster) %>% top_n(20, avg_logFC)
```

### Assign global identities


Assign identify based on canonical markers

```{r fig.height=8, fig.width=8}

DimPlot(int, label = T)
DimPlot(int, group.by = "Phase")

FeaturePlot(int, features = c("PDGFRA", "PAX2", "SOX17", "PECAM1"), order = T)

FeaturePlot(int, features = c("PAX2", "PAX8", "LHX1", "JAG1"), order = T)

FeaturePlot(int, features = c("CITED1", "SIX2", "ROBO2", "WT1"), order = T)

FeaturePlot(int, features = c("MEIS1", "SIX1","SOX9","PAX2"))

FeaturePlot(int, features = c("GATA3", "HNF4A","SLC12A1","MAFB"), order = T)

ggsave(filename = "plots/all/global-markers.png",
FeaturePlot(int, features = c("WT1","PAX2","PDGFRA","SOX17"), ncol = 1, coord.fixed = T),
width = 4, height = 14)
```



```{r assign-global-cell-type-and-plot}

nephron.cluster <- c(4,11,10,12)
endothelial.cluster <- c(14)
total.clusters <- unique(int@active.ident)

#assign nephron lineage clusters
int@meta.data['neph'] <- int@active.ident %in% nephron.cluster

#assign endothelial clusters
int@meta.data['endo'] <- int@active.ident %in% endothelial.cluster

#assign stomral clusters
int@meta.data['stroma'] <- int@active.ident %in% (setdiff(total.clusters, c(nephron.cluster, endothelial.cluster)))


int@meta.data['global.type'] <- NA
int@meta.data$global.type[int@meta.data$neph] <- 'Nephron'
int@meta.data$global.type[int@meta.data$stroma] <- 'Stroma'
int@meta.data$global.type[int@meta.data$endo] <- 'Endothelial'
int@meta.data$global.type <- factor(int@meta.data$global.type, levels = c("Stroma","Nephron","Endothelial"))

summary(int@meta.data$global.type)

global.type.colours <- c("orange","light blue","dark grey")

global.type.umap <- DimPlot(int, group.by = 'global.type', 
                            label = F, 
                            cols = global.type.colours) 

ggsave(filename = "plots/all/global-type-umap.png", units = "mm", width = 150, height = 120)
```
```{r global-plots-supp-figure, fig.height=8, fig.width=12}
p1 <- DimPlot(int, label = T)
p2 <- DimPlot(int, group.by = "Phase")
p3 <- DimPlot(int, group.by = "global.type", label = TRUE)
p4 <- DimPlot(int, group.by = "orig.ident")


plt <- gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

ggsave(filename = "plots/all/global-clusters-umap.png",
       plot = plt, width = 12, height = 8)
```




```{r plotting-global-proportions}

global.props <- as.data.frame(prop.table(table(int@meta.data$rep, int@meta.data$global.type), margin = 1))

colnames(global.props) <- c("replicate","cluster","prop")

global.props["condition"] <- sapply(as.character(global.props$replicate), 
                      FUN = function(x){ return(strsplit(x, split = "_")[[1]][1]) } )

global.props <- global.props %>% group_by(cluster) %>% dplyr::mutate('norm.prop' = prop / mean(prop))

global.ratio.plot <- ggplot(global.props, aes(x =condition, y = prop*100)) + 
  geom_point() +
  stat_summary(fun.y = "mean", 
               geom = "point", 
               size = 5, 
               shape = 18, 
               colour = "red", 
               alpha = 0.5) + 
  facet_wrap(~cluster, ncol = 1, scales = "free") + 
   theme_classic() +  
  xlab(label = "") +
  ylab(label = "%") +
    theme(strip.text.x = element_text(size=12), 
          legend.title = element_blank(), 
          legend.text = element_text(size = 10))
  
ggsave(global.ratio.plot, filename = "plots/all/global_ratio_plot.png", dpi = 300, width = 50, height = 120, units = 'mm')


```

statistical testing for cluster proportions

```{r}
for (c in unique(global.props$cluster)){
st <- global.props %>% dplyr::filter(cluster == c)
aa <- aov(prop~condition, data = st)
print(paste("cluster", c))
print(summary(aa))
print(TukeyHSD(aa))
print(pairwise.t.test(st$prop, st$condition))
}
```




### Nephron sub-clustering

```{r}

if(!file.exists(paste0("output/integrated-", condition.string, "-neph-subclust.RDS"))){
  int.neph <- subset(int, subset = global.type == 'Nephron')
  int.neph <- sub_cluster(int.neph)
  saveRDS(int.neph, paste0("output/integrated-", condition.string, "-neph-subclust.RDS"))
}else{
  int.neph <- readRDS(paste0("output/integrated-", condition.string, "-neph-subclust.RDS"))
}

DefaultAssay(int.neph) <- "RNA"
int.neph <- NormalizeData(int.neph)

```

```{r clustree-neph, fig.height=6, fig.width=8}
clustree(int.neph, assay = "SCT")
```

```{r assign-resolution-UMAP}
int.neph <- SetIdent(int.neph, value = int.neph@meta.data$SCT_snn_res.0.6)

DimPlot(int.neph, label = T)
```

```{r QC-metrics, fig.height=8, fig.width=8}

p1 <- VlnPlot(int.neph, features = "nCount_RNA", pt.size = 0)
p2 <- VlnPlot(int.neph, features = "nFeature_RNA", pt.size = 0)
p3 <- VlnPlot(int.neph, features = "percent.mito", pt.size = 0)
p4 <- VlnPlot(int.neph, features = "scrublet_score", pt.size = 0)

gridExtra::grid.arrange(p1, p2, p3, p4)
```


```{r examine-doublets, fig.height=4, fig.width=8}
clust.plt <- DimPlot(int.neph, label = T)
phase.plt <- DimPlot(int.neph, group.by = "Phase")
dbl.plt <- FeaturePlot(int.neph, features = "scrublet_score")
set.plt <- DimPlot(int.neph, group.by = "orig.ident")

gridExtra::grid.arrange(clust.plt, phase.plt, dbl.plt, set.plt, ncol = 2)

```


```{r remove-doublets}
#remove cluster 6 and 9 - see above

int.neph <- subset(int.neph, idents = c(0,1,2,3,4,5,7,8))

```

Plotting various canonical marker genes

```{r}
FeaturePlot(int.neph, features = c("PAX2", "PAX8", "LHX1", "JAG1"), order = T)

FeaturePlot(int.neph, features = c("CITED1", "SIX2", "ROBO2", "WT1"), order = T)

FeaturePlot(int.neph, features = c("PDGFRA", "DCN","PBX1","FOXD1"), order = T)

FeaturePlot(int.neph, features = c("GATA3", "HNF4A","SLC12A1","MAFB"), order = T)

FeaturePlot(int.neph, features = c("NPHS2", "PODXL","CTGF","MAFB"), order = T)
```

Feature plot for figure

```{r plot-nephron-mafb-gata3-hnf4a}

plt <- FeaturePlot(int.neph, features = c("MAFB", "HNF4A", "GATA3"), ncol = 1, coord.fixed = T)

ggsave(filename = "plots/all/neph-mafb-hnf4a-gata3.png", plot = plt, dpi = 300, units = "in", width = 4, height = 4)

```

Umap plot for figure

```{r nephron-umap-plot, fig.height=3, fig.width=4}

neph.cluster.plt <- DimPlot(int.neph, label = T) 

ggsave(filename = "plots/all/neph_UMAP.png", plot = neph.cluster.plt, 
       units = "in", width = 4, height = 3)

DimPlot(int.neph, label = T)

```

More plots for figures

```{r id-plots, fig.height=9, fig.width=12}
ggsave(filename = "plots/all/neph-markers-for-supp.png",
FeaturePlot(int.neph, features = c("GATA3","HNF1B","CUBN","HNF4A",
                                   "FOXC2","MAFB","PODXL","SIX2","EYA1"), ncol = 3, coord.fixed = T),
width = 12, height = 9 
)

```

Plotting predicted phase for supp figures

```{r fig.height=6, fig.width=8}
clust.plt <- DimPlot(int.neph, label = T) 
phase.plt <- DimPlot(int.neph, group.by = "Phase")
set.plt <- DimPlot(int.neph, group.by = "orig.ident")

ggsave(filename = "plots/all/neph-clust-phase-for-supp.png",
       gridExtra::grid.arrange(clust.plt, phase.plt, set.plt, ncol = 2),
       width = 8, height = 6)
```

Plotting proportions for figure

```{r plotting-proportions, fig.height=1.5, fig.width=6}

# get proportions, excluding cycling cells

neph.props <- get_props_table(int.neph)

neph.props$condition <- factor(neph.props$condition, 
                               levels = c("Man","R0","R40"))

clusters.to.plot <- c(0,1,2,3,4,8)

neph.props.forplot <- neph.props %>% dplyr::filter(cluster %in% clusters.to.plot)

neph.props.forplot['named.cluster'] <- factor(neph.props.forplot$cluster, 
              levels = c(3,4,1,2,0,8),
              labels = c("Progenitor (3)", "Pre-Pod. (4)", "Podoctye (1)", "Pre-Tubule (2)","Distal (0)","Proximal (8)"))

neph.props.plt <- ggplot(neph.props.forplot,
                         aes(x = condition, y = prop*100)) + 
  geom_point(size =1) + 
  stat_summary(fun.y = "mean", geom = "point", shape=18, size = 4, colour = "red", alpha = 0.5)+
  facet_wrap(~named.cluster, scales = "free", nrow=1) +
  theme_classic() +
  xlab(label = "") +
  ylab(label = "%") +
    theme(axis.line.x.bottom = element_blank(), axis.ticks.x = element_blank(),
          strip.text.x = element_text(size=8), strip.background = element_blank(), 
          axis.text = element_text(size=8))

ggsave(neph.props.plt, filename = "plots/all/neph_clust_prop.png", dpi = 300,
        units = "in", height = 1.5, width = 6)


```


Statistical testing for nephron proportions

```{r stats-testing-props}

for (c in unique(neph.props$cluster)){
st <- neph.props %>% dplyr::filter(cluster == c)
aa <- aov(prop~condition, data = st)
print(paste("cluster", c))
print(summary(aa))
print(TukeyHSD(aa))
print(pairwise.t.test(st$prop, st$condition))
}

```


Testing DE expression within cluster :


```{r nephron-DE-testing-all-clusters}
for (c in unique(int.neph@active.ident)){
  print(paste0("cluster = ", c))
  de.glm <- do_pairwise_edgeR(subset(int.neph, idents = c))
  out.path <- paste0("DE-testing/neph_cluster-", c, "_")
  print(out.path)
  write.csv(x = de.glm$r40_man$table, file = paste0(out.path, "R40_v_man.csv"))
  write.csv(x = de.glm$r40_r0$table, file = paste0(out.path, "R40_v_r0.csv"))
  write.csv(x = de.glm$r0_man$table, file = paste0(out.path, "R0_v_man.csv"))
  write.csv(x = de.glm$any$table, file = paste0(out.path, "any.csv"))
}

```

Filtering DE testing to remove genes that are expressed in more than 3 clusters

```{r}

all_r40_r0 <- lapply(unique(int.neph@active.ident), function(x) {
  de <- do_pairwise_edgeR(subset(int.neph, idents = x))
  out <- de$r40_r0$table
  out['gene'] <- rownames(out)
  out['cluster'] <- x
  return(out)
})
all_r40_r0 <- do.call(rbind, all_r40_r0)
write.csv(all_r40_r0, "DE-testing/all_r40_r0-DE.csv")
all_r40_r0 %>% group_by(cluster) %>% summarise("n.de" = n())
all_r40_r0 %>% group_by(gene) %>% filter(n()>2) %>% summarise('count' = n())%>% arrange(-count)
all_r40_r0 %>% group_by(gene) %>% filter(n()< 4) %>% write.csv(file = "DE-testing/neph-filtered-under4-r40_r0.csv")


all_r40_man <- lapply(unique(int.neph@active.ident), function(x) {
  de <- do_pairwise_edgeR(subset(int.neph, idents = x))
  out <- de$r40_man$table
  out['gene'] <- rownames(out)
  out['cluster'] <- x
  return(out)
})

all_r40_man <- do.call(rbind, all_r40_man)
write.csv(all_r40_man, "DE-testing/all_r40_man-DE.csv")
all_r40_man %>% group_by(cluster) %>% summarise("n.de" = n())
all_r40_man %>% group_by(gene) %>% filter(n()>2) %>% summarise('count' = n()) %>% arrange(-count)
all_r40_man %>% group_by(gene) %>% filter(n()< 4) %>% write.csv(file = "DE-testing/neph-filtered-under4-r40_man.csv")

all_any <- lapply(unique(int.neph@active.ident), function(x) {
  de <- do_pairwise_edgeR(subset(int.neph, idents = x))
  out <- de$any$table
  out['gene'] <- rownames(out)
  out['cluster'] <- x
  return(out)
})

all_any <- do.call(rbind, all_any)
write.csv(all_any, "DE-testing/all_any-DE.csv")
all_any %>% group_by(cluster) %>% summarise("n.de" = n())
all_any %>% group_by(gene) %>% filter(n()>2) %>% summarise('count' = n()) %>% arrange(-count)
all_any %>% group_by(gene) %>% filter(n()< 4) %>% write.csv(file = "DE-testing/neph-filtered-under4-any.csv")

all_r0_man <- lapply(unique(int.neph@active.ident), function(x) {
  de <- do_pairwise_edgeR(subset(int.neph, idents = x))
  out <- de$r0_man$table
  out['gene'] <- rownames(out)
  out['cluster'] <- x
  return(out)
})

all_r0_man <- do.call(rbind, all_r0_man)
write.csv(all_r0_man, "DE-testing/all_r0_man-DE.csv")
all_r0_man %>% group_by(cluster) %>% summarise("n.de" = n())
all_r0_man %>% group_by(gene) %>% filter(n()>2) %>% summarise('count' = n()) %>% arrange(-count)
all_r0_man %>% group_by(gene) %>% filter(n()< 4) %>% write.csv(file = "DE-testing/neph-filtered-under4-r0_man.csv")

```


Generating tables and plots of DE testing results

```{r}

r40_r0.table <- all_r40_r0 %>% group_by(gene) %>% filter(n()< 4) %>% group_by(cluster) %>% summarise(n.de = n())


r40_man.table <- all_r40_man %>% group_by(gene) %>% filter(n()< 4) %>% group_by(cluster) %>% summarise(n.de = n())
write.csv(file = "DE-testing/neph-filtered-under4-r40_man.csv", x = r40_man.table)

r0_man.table <- all_r0_man %>% group_by(gene) %>% filter(n()< 4) %>% group_by(cluster) %>% summarise(n.de = n())
write.csv(file = "DE-testing/neph-filtered-under-4-r0_man.csv", x = r0_man.table)

de.table <- dplyr::full_join(r40_r0.table, r40_man.table, by = "cluster") %>% as.data.frame()
colnames(de.table) <- c("cluster","R40-R0", "R40-Man")

de.table <- dplyr::full_join(de.table, r0_man.table, by = "cluster") %>% as.data.frame()
colnames(de.table)[4] <- "R0-Man"

de.table[is.na(de.table)] <- 0

rownames(de.table) <- labels(de.table$cluster)

de.table <- de.table %>% filter(cluster %in% c(0,1,2,3,4,8))

de.table['cell.type'] <- c("Distal","Podocyte","Mesenchyme","Pre-Podocyte","Pre-Tubule","Proximal")

pdf("plots/all/DE-genes-per-clust.pdf", width = 6, height = 4)
gplots::heatmap.2(as.matrix(de.table[,2:4]), Colv = TRUE, labRow = de.table$cell.type, 
                  Rowv = TRUE, dendrogram = "none", col = viridis::viridis(60), 
                  density.info = "none", trace = 'none', cexRow = 0.5, cexCol = 0.5,  
                  keysize = 1,
                  key.par = list(cex=0.5), key.title = "", key.xlab = "No. DE genes")
dev.off()



```


Plotting proximal DE genes of interest for figure

```{r vln-plot-for-fig, fig.height=3, fig.width=9}

prox.list <- c("SLC30A1","SLC51B","FABP3","SULT1E1","SPP1", "JAG1")

cc <- subset(int.neph, idents = 8)

cc <- SetIdent(cc, value = cc$orig.ident)

prox.plt <- VlnPlot(cc, features = prox.list, ncol = 6, pt.size = 0.5)

ggsave(filename = "plots/all/proximal-for-fig.png", prox.plt, width = 9, height = 3) 

prox.plt
```


Marker genes for nephron cluster

```{r nephron-markers}

neph.marker.file <- paste0('output/', condition.string, '-neph-cluster-markers.RDS')

if(!file.exists(neph.marker.file)){
  markers <- FindAllMarkers(int.neph, logfc.threshold = 0.25, only.pos = TRUE)
  saveRDS(markers, neph.marker.file)
  export_marker_xls(markers, xls.path = paste0("output/", condition.string, "-neph_cluster_markers.xlsx"))
}

neph.markers <- readRDS(neph.marker.file)
```

```{r}
neph.markers %>% group_by(cluster) %>% top_n(10, pct.1-pct.2)

```


Comparison to cluster markers from Combes, Zappia et al

```{r compare-combes-zappia}

cz.pod <- c("TCF21", "PODXL", "VEGFA", "NPHS1", "PTPRO")

FeaturePlot(int.neph, features = cz.pod, ncol = 3)

cz.pre.pod <- c("CTGF", "OLFM3", "MAFB", "NPHS1")

FeaturePlot(int.neph, features = cz.pre.pod)

cz.prox.pre <- c("IGFBP7", "FXYD2", "CDH6", "HNF1B")

FeaturePlot(int.neph, features = cz.prox.pre)

cz.distal.pre <- c("EPCAM", "EMX2", "SPP1", "MAL", "PAX2")

FeaturePlot(int.neph, features = cz.distal.pre, ncol = 3)

cz.progenitor <- c("DAPL1", "LYPD1", "SIX1", "CRABP2")

FeaturePlot(int.neph, cz.progenitor)

```



### Stromal sub-clustering


```{r}

stroma.cluster.file <- paste0("output/integrated-", condition.string, "-stroma-subclust.RDS")

if(!file.exists(stroma.cluster.file)){
  int.stroma <- subset(int, subset = global.type == 'Stroma')
  int.stroma <- sub_cluster(int.stroma)
  saveRDS(int.stroma, stroma.cluster.file)
}else{
  int.stroma <- readRDS(stroma.cluster.file)
}

DefaultAssay(int.stroma) <- "RNA"
int.stroma <- NormalizeData(int.stroma)


```


```{r fig.width=10}
clustree(int.stroma, assay = "SCT")
```


```{r stroma-umap, fig.height=10, fig.width=4}
int.stroma <- SetIdent(int.stroma, value = int.stroma@meta.data$SCT_snn_res.0.4)

int.stroma@active.ident <- factor(int.stroma@active.ident, levels = seq(0,10))

p1 <- DimPlot(int.stroma, label = T)
p2 <- DimPlot(int.stroma, group.by = "Phase")
p3 <- DimPlot(int.stroma, group.by = "orig.ident")

pp <- gridExtra::grid.arrange(p1, p2, p3, ncol = 1)

ggsave(filename = "plots/all/stroma-UMAP-clusters-phase-set.png", plot = pp, width = 4, height = 10)

```


```{r qc-plots, fig.height=12, fig.width=10}

p1 <- VlnPlot(int.stroma, features = "nCount_RNA")
p2 <- VlnPlot(int.stroma, features = "nFeature_RNA")
p3 <- VlnPlot(int.stroma, features = "percent.mito")
p4 <- VlnPlot(int.stroma, features = "scrublet_score", pt.size = 0)

gridExtra::grid.arrange(p1, p2, p3, p4)


```

Get stromal proportions and plot

```{r plot-stroma-props, fig.height=5, fig.width=2}
stroma.props <- get_props_table(int.stroma)


stroma.props$cluster <- factor(stroma.props$cluster, levels = c(seq(0,10)))

stroma.props.plt <- stroma.props %>% filter(!(cluster %in% c(4,5,6)))

stromal.prop.plt <- ggplot(stroma.props.plt, aes(x = condition, y = prop * 100)) + 
  geom_point(size = 2) + 
  stat_summary(fun.y = "mean", geom = "point", shape=18, size = 5, colour = "red", alpha = 0.5)+
  facet_wrap(~cluster, scales = "free", ncol=2) +
  theme_classic() +
  xlab(label = "") +
  ylab(label = "%") +
    theme(axis.line.x.bottom = element_blank(), axis.ticks.x = element_blank(),
          strip.text.x = element_text(size=14), strip.background = element_blank(), 
          axis.text = element_text(size=14))

ggsave(filename = "plots/all/stroma-proportions.png", stromal.prop.plt, width = 4, height = 6)


```


Statistical testing for stromal cluster proportions

```{r stats-test-stroma}

for (c in unique(stroma.props$cluster)){
st <- stroma.props %>% dplyr::filter(cluster == c)
aa <- aov(prop~condition, data = st)
print(paste("cluster", c))
print(summary(aa))
print(TukeyHSD(aa))
}


```

Generate stromal marker lists

```{r stromal-markers}

stroma.marker.file <- paste0('output/', condition.string, '-stroma-cluster-markers.RDS')

if(!file.exists(stroma.marker.file)){
  stroma.markers <- FindAllMarkers(int.stroma, logfc.threshold = 0.25, only.pos = TRUE)
  saveRDS(stroma.markers, stroma.marker.file)
  export_marker_xls(stroma.markers, xls.path = paste0("output/", condition.string, "-stroma_cluster_markers.xlsx"))
} else {
stroma.markers <- readRDS(stroma.marker.file)}
```


```{r}
stroma.markers %>% group_by(cluster) %>% top_n(40, pct.1-pct.2)

```


Plotting interesting cluster : 7, 10

```{r fig.height=6, fig.width=12}

#plotting clusters with changes in proportion - 7, 10

s.plot.1 <- FeaturePlot(int.stroma, features = c("COL2A1","LHX9","S100A10","WNT5A",
                                     "ZIC1","ZIC4","ANGPTL1","SEMA5A"), ncol = 4, coord.fixed = T)

ggsave(filename = "plots/all/stromal-feature-plot-c7-10.png", plot = s.plot.1, width = 12, height = 6)

```

Plotting cluster markers for extended data figures

```{r fig.height=6, fig.width=6}

plt <- FeaturePlot(int.stroma, features = c("CITED1","SIX2","GATA","FOXC2"), ncol = 2, coord.fixed = T)

ggsave(filename = "plots/all/stroma-clust-2-markers.png",plot = plt)

```

Plotting stromal markers for extended data imaging supp.

```{r plt-for-imaging-supp}

ggsave(filename = "plots/all/stroma-IF-markers.png",
       plot = FeaturePlot(int.stroma, features = c("MEIS1","MEIS2","SIX1","SOX9"), ncol = 4, coord.fixed = T), width = 14)


```

Plotting genes for figure

```{r clust-2-vln-for-fig, fig.height=4, fig.width=3}

ggsave(filename = "plots/all/stroma-clust2-vln.png", 
plot_by_set(int.stroma, clust = 2, genes = c("RSPO3","WNT5A"), ncol = 2, pt.size = 0.01),
width = 4, height = 3)


```

Plottings genes for figure

```{r clust3-vln-for-fig, fig.height=3, fig.width=8}

ggsave(filename = "plots/all/stroma-clust3-vln.png", 
plot_by_set(int.stroma, clust = 3, genes = c("HOXA11","FOXC2","EYA1","SIX1"), ncol = 4, pt.size = 0.01),
width = 8, height = 3)

```


Plotting for extended data figure

```{r stroma-plot-supp-fig, fig.height=10, fig.width=6}

ggsave(filename = "plots/all/stroma-clust-markers-supp.png",
FeaturePlot(int.stroma, features = c("SIX2","LYPD1","FOXC2","HOXA11",
                                     "WNT5A","LHX9",
                                     "ZIC1","ZIC4"), ncol = 2),
width = 6, height = 10)


```

DE testing within stromal clusters

```{r stroma-cluster-de}

get_stromal_de <- function(obj){

all_r40_r0 <- lapply(unique(obj@active.ident), function(x) {
  de <- do_pairwise_edgeR(subset(obj, idents = x))
  out <- de$r40_r0$table
  out['gene'] <- rownames(out)
  out['cluster'] <- x
  return(out)
})
all_r40_r0 <- do.call(rbind, all_r40_r0)
write.csv(all_r40_r0, "DE-testing/stroma/stroma_r40_r0-DE.csv")
all_r40_r0 %>% group_by(cluster) %>% summarise("n.de" = n())
all_r40_r0 %>% group_by(gene) %>% filter(n()>2) %>% summarise('count' = n())%>% arrange(-count)


all_r40_man <- lapply(unique(obj@active.ident), function(x) {
  de <- do_pairwise_edgeR(subset(obj, idents = x))
  out <- de$r40_man$table
  out['gene'] <- rownames(out)
  out['cluster'] <- x
  return(out)
})

all_r40_man <- do.call(rbind, all_r40_man)
write.csv(all_r40_man, "DE-testing/stroma/stroma_r40_man-DE.csv")
all_r40_man %>% group_by(cluster) %>% summarise("n.de" = n())
all_r40_man %>% group_by(gene) %>% filter(n()>2) %>% summarise('count' = n()) %>% arrange(-count)


all_any <- lapply(unique(obj@active.ident), function(x) {
  de <- do_pairwise_edgeR(subset(obj, idents = x))
  out <- de$any$table
  out['gene'] <- rownames(out)
  out['cluster'] <- x
  return(out)
})

all_any <- do.call(rbind, all_any)
write.csv(all_any, "DE-testing/stroma/stroma_any-DE.csv")
all_any %>% group_by(cluster) %>% summarise("n.de" = n())
all_any %>% group_by(gene) %>% filter(n()>2) %>% summarise('count' = n()) %>% arrange(-count)


all_r0_man <- lapply(unique(obj@active.ident), function(x) {
  de <- do_pairwise_edgeR(subset(obj, idents = x))
  out <- de$r0_man$table
  out['gene'] <- rownames(out)
  out['cluster'] <- x
  return(out)
})

all_r0_man <- do.call(rbind, all_r0_man)
write.csv(all_r0_man, "DE-testing/stroma/stroma_r0_man-DE.csv")
all_r0_man %>% group_by(cluster) %>% summarise("n.de" = n())
all_r0_man %>% group_by(gene) %>% filter(n()>2) %>% summarise('count' = n()) %>% arrange(-count)

r40_r0.table <- all_r40_r0 %>% group_by(gene) %>% filter(n()< 4) %>% group_by(cluster) %>% summarise(n.de = n())

r40_man.table <- all_r40_man %>% group_by(gene) %>% filter(n()< 4) %>% group_by(cluster) %>% summarise(n.de = n())

r0_man.table <- all_r0_man %>% group_by(gene) %>% filter(n()< 4) %>% group_by(cluster) %>% summarise(n.de = n())

de.table <- dplyr::full_join(r40_r0.table, r40_man.table, by = "cluster") %>% as.data.frame()
colnames(de.table) <- c("cluster","R40-R0", "R40-Man")

de.table <- dplyr::full_join(de.table, r0_man.table, by = "cluster") %>% as.data.frame()
colnames(de.table)[4] <- "R0-Man"

de.table[is.na(de.table)] <- 0

rownames(de.table) <- labels(de.table$cluster)

return(de.table)

}

stromal.de.table <- get_stromal_de(int.stroma)

```



Generate summary table of stromal DE genes

```{r fig.height=6, fig.width=6}

plt.table <- stromal.de.table %>% filter(!(cluster %in% c(4,5,6)))

pdf("plots/all/stromal-DE-genes-per-clust.pdf", width = 6, height = 6)
gplots::heatmap.2(as.matrix(plt.table[,2:4]), Colv = TRUE, labRow = plt.table$cluster, 
                  Rowv = TRUE, dendrogram = "none", col = viridis::viridis(60), 
                  density.info = "none", trace = 'none', cexRow = 2, cexCol = 2,  
                  keysize = 1, margins = c(10,10),
                  key.par = list(cex=0.5), key.title = "", key.xlab = "No. DE genes")
dev.off()


```

Generate filtered stromal DE lists


```{r fig.height=4, fig.width=8}

df <- read.csv("DE-testing/stroma/stroma_r40_man-DE.csv", stringsAsFactors = F)
df <- df %>% group_by(gene) %>% filter(n()<4) %>% ungroup() %>% group_by(cluster) 
write.csv(file = "DE-testing/stroma/filtered_4-or-less-r40_man-DE.csv", x = df)

df <- read.csv("DE-testing/stroma/stroma_r40_r0-DE.csv", stringsAsFactors = F)
df <- df %>% group_by(gene) %>% filter(n()<4) %>% ungroup() %>% group_by(cluster) 
write.csv(file = "DE-testing/stroma/filtered_4-or-less-r40_r0-DE.csv", x = df)

df <- read.csv("DE-testing/stroma/stroma_r0_man-DE.csv", stringsAsFactors = F)
df <- df %>% group_by(gene) %>% filter(n()<4) %>% ungroup() %>% group_by(cluster) 
write.csv(file = "DE-testing/stroma/filtered_4-or-less-r0_man-DE.csv", x = df)

```



### bulk-comparison

This is a comparison to bulk RNAseq data.


```{r load-bulk-data}

## note - need to change path below to correct location

bulk.de <- read.csv("/group/kidn1/hpc/KEG/Profiling_RNAseq/DD104_MAFB-GATA3_bulkRNAseq/analysis/out/R40_v_R0_all_DE.csv", stringsAsFactors = F)

sig.de <- bulk.de %>% dplyr::filter(PValue < 0.05)

colnames(sig.de)[1] <- "gene"
rownames(sig.de) <- sig.de$gene
  
top100 <- sig.de %>% dplyr::top_n(-100, PValue) %>% arrange(logFC)


```

Generate heatmap comparison top bulk DE genes to pseudo-bulk for clusters.

```{r fig.height=12, fig.width=12}

clust.sce <- as.SingleCellExperiment(SetIdent(int, value = int@meta.data$global.type))
clust.agg <- scater::sumCountsAcrossCells(clust.sce,
               ids=paste0(clust.sce@colData$rep, clust.sce@colData$global.type))
               
colnames(clust.agg)

logcounts <- cpm(as.matrix(clust.agg), log = T)

top100.common <- intersect(rownames(logcounts),top100$gene)

logcounts[top100.common,]

pdf(file = "plots/all/DE-cluster-comparison.pdf", width = 12, height = 12)
gplots::heatmap.2(logcounts[top100.common,], trace = "none", scale = "row", 
                  col = viridis::viridis(60), margins = c(10,10), 
                  keysize = 1, lhei = c(1,10), key.title = NA)
dev.off()

```



```{r}
sessionInfo()
```


