---
title: "Quality control of Bioprint Datasets 2020"
output:
  html_document:
    df_print: paged
---

This is a record of analysis of three datasets generated from kidney organoids with three different conformations :

'hand' - traditional hand made organoids using a centrifuge. In the manuscript these are refereed to as 'manual' organoids.

'dots' - single deposition bioprinted organonids to replicate the handmade protocol. In the manuscript these are referred to as 'R0' organoids.

'line' - line organoids, where the same number of cells as 'dots' are spread out over a thing line with movement:extrusion ration of 40:1 (distance moved : syringe plunger displacement). In the manuscript these are referred to as 'R40' organoids.


This notebook has been designed to be run either in Rstudio, or on a HPC. In the latter case multicore processing is enabled. In Rstudio, default behaviour should be single core.

```{r load-libraries}
library(Seurat)
library(ggplot2)
library(future)
library(dplyr)
library(clustree)

if (supportsMulticore()){
  
plan('multiprocess', workers=8)
options(future.globals.maxSize = 12000 * 1024^2)
print('Using parallel processing')

} else {
  print('multi-core disabled on Rstudio')
}


```



```{r load-data}

if(!file.exists("output/load-all.RDS")){

load_data <- function(location, project_name){
  seurat <- Read10X(paste0(location, "outs/filtered_feature_bc_matrix/"))
  seurat <- CreateSeuratObject(counts = seurat, project = project_name, min.cells = 20)
  seurat <- PercentageFeatureSet(seurat, pattern = "^MT-", col.name = "percent.mito")
  
  scrublet <- read.csv(paste0(location, "outs/scrublet/predicted_doublets.csv"))
  seurat@meta.data['scrublet_score'] <- scrublet$doublet_scores
  seurat@meta.data['scrublet_prediction'] <- scrublet$predicted_doublets
  
  return(seurat)
}

dt <- load_data("../dots/BP-dots/", 'dots')
ln <- load_data("../line/BP-line/", 'line')
hd <- load_data("../hand/BP-hand/", 'hand')

sets.list <- list(dt, ln, hd)

saveRDS(object=sets.list, file="output/load-all.RDS")

}else
  {
  sets.list <- readRDS("output/load-all.RDS")
}


```

Define function to load Hash Tag Oligo (HTO) data

```{r}
add_hto <- function(cdna, hto_count_path, cutoff){

hto <- Read10X(data.dir = hto_count_path, gene.column = 1)

#remove the 'unmapped' row as this confuses the automatic HTO assignment below
oligos <- rownames(hto)[1:4]
hto <- hto[oligos,]

#use only matching barcodes
cdna.barcodes <- colnames(cdna)
hto.barcodes <- colnames(hto)
common.barcodes <- intersect(hto.barcodes, cdna.barcodes)
cdna <- cdna[,common.barcodes]
hto <- hto[,common.barcodes]

#add HTO data, normalise and scale
cdna[["HTO"]] <- CreateAssayObject(counts = hto)
cdna <- NormalizeData(cdna, assay = "HTO", normalization.method = "CLR")
cdna <- ScaleData(cdna, assay = "HTO")
cdna <- HTODemux(cdna, assay = 'HTO', positive.quantile = cutoff, init = 5)

return(cdna)
}

```



```{r hto-load-dots, fig.height=12}

dots <- sets.list[[1]]
dots <- add_hto(dots, '../dots/cite-seq-count/umi_count/', 0.993)
table(dots@meta.data$HTO_classification.global)
table(dots@meta.data$hash.ID)
RidgePlot(dots, features = unique(dots@meta.data$hash.ID), group.by = "hash.ID", ncol = 1)

```



```{r hto-load-line, fig.height=12}

line <- sets.list[[2]]
line <- add_hto(line, '../line/cite-seq-count/umi_count/', 0.90)
table(line@meta.data$HTO_classification.global)
table(line@meta.data$hash.ID)
RidgePlot(line, features = unique(line@meta.data$hash.ID), group.by = "hash.ID", ncol = 1)

```

```{r hto-load-hand, fig.height=12}

hand <- sets.list[[3]]
hand <- add_hto(hand, '../hand/cite-seq-count/umi_count/', 0.999)
table(hand@meta.data$HTO_classification.global)
table(hand@meta.data$hash.ID)
RidgePlot(hand, features = unique(hand@meta.data$hash.ID), group.by = "hash.ID", ncol = 1)

```




Visualisation for each dataset (slow to run)


```{r}

hto.dist.mtx <- as.matrix(dist(t(GetAssayData(object = dots, assay = "HTO"))))
dots <- RunTSNE(dots, distance.matrix = hto.dist.mtx, perplexity = 50)
DimPlot(dots, group.by = 'hash.ID', reduction = "tsne")

```


```{r}

# this takes ages to run :

hto.dist.mtx <- as.matrix(dist(t(GetAssayData(object = line, assay = "HTO"))))
line <- RunTSNE(line, distance.matrix = hto.dist.mtx, perplexity = 50)
DimPlot(line, group.by = 'hash.ID', reduction = "tsne")

```

```{r}
hto.dist.mtx <- as.matrix(dist(t(GetAssayData(object = hand, assay = "HTO"))))
hand <- RunTSNE(hand, distance.matrix = hto.dist.mtx, perplexity = 50)
DimPlot(hand, group.by = 'hash.ID', reduction = "tsne")
```



```{r remove-doublets}

keep_singlets <- function(seurat){
  seurat <- SetIdent(seurat, value = seurat@meta.data$HTO_classification.global)
  seurat <- subset(seurat, idents = "Singlet")
  return(seurat)
}



dots <- keep_singlets(dots)
line <- keep_singlets(line)
hand <- keep_singlets(hand)

```


Plotting for QC

```{r fig.height=15, fig.width=10}

p1 <- FeatureScatter(line, feature1 = "nCount_RNA", feature2 = 'percent.mito') + geom_hline(yintercept =15)
p2 <- FeatureScatter(dots, feature1 = "nCount_RNA", feature2 = 'percent.mito') + geom_hline(yintercept =15)
p3 <- FeatureScatter(hand, feature1 = "nCount_RNA", feature2 = 'percent.mito')+ geom_hline(yintercept =15)


gridExtra::grid.arrange(p1, p2, p3, ncol=2)

```

Filter cells with greater than 15% mitochondrial content as a metric of cell health.

```{r subset}
sets.list <- list(dots, line, hand)

filter_sets <- function(set){
  set <- subset(set, subset = percent.mito < 15 )
  return(set)
}

sets.list <- lapply(sets.list, filter_sets)

```



```{r fig.height=15, fig.width=10}

cutoff = 1000

p1 <- FeatureScatter(sets.list[[1]], feature1 = "percent.mito", feature2 = 'nFeature_RNA') + scale_y_log10() + geom_hline(yintercept =cutoff)

p2 <- FeatureScatter(sets.list[[2]], feature1 = "percent.mito", feature2 = 'nFeature_RNA') + scale_y_log10() +
  geom_hline(yintercept =cutoff )
p3 <- FeatureScatter(sets.list[[3]], feature1 = "percent.mito", feature2 = 'nFeature_RNA') + scale_y_log10() +
  geom_hline(yintercept =cutoff) 


gridExtra::grid.arrange(p1, p2, p3, ncol=2)

```


```{r}
filter_sets <- function(set){
  set <- subset(set, subset = nFeature_RNA > 1000 )
  return(set)
}

sets.list <- lapply(sets.list, filter_sets)

```

Now Looking at doublet detection scores from scrublet

```{r fig.height=12, fig.width=10}
p1 <- FeatureScatter(sets.list[[1]], feature1 = "scrublet_score", feature2 = 'nFeature_RNA')
p2 <- FeatureScatter(sets.list[[2]], feature1 = "scrublet_score", feature2 = 'nFeature_RNA') 
p3 <- FeatureScatter(sets.list[[3]], feature1 = "scrublet_score", feature2 = 'nFeature_RNA') 


gridExtra::grid.arrange(p1, p2, p3, ncol=2)
```

```{r fig.height=12, fig.width=10}
p1 <- VlnPlot(sets.list[[1]], features = "scrublet_score", y.max = 1)
p2 <- VlnPlot(sets.list[[2]], features = "scrublet_score", y.max = 1)
p3 <- VlnPlot(sets.list[[3]], features = "scrublet_score", y.max = 1)

gridExtra::grid.arrange(p1, p2, p3, ncol=3)

```

Check details

```{r}
sets.list[[1]]

sets.list[[2]]

sets.list[[3]]
```



Run SCtransform here

```{r}

sctransform_all <- function(set){
  set <- SCTransform(set)
  return(set)
}

sets.list <- parallel::mclapply(sets.list, sctransform_all, mc.cores = 4)

saveRDS(sets.list, "output/sctransform-all.RDS")

```

Integration

```{r}

integrate_sets <- function(sets.list){
  
  set.features <- SelectIntegrationFeatures(object.list = sets.list, nfeatures = 3000)
  sets.list <- PrepSCTIntegration(object.list = sets.list, anchor.features = set.features)
  sets.anchors <- FindIntegrationAnchors(object.list = sets.list, normalization.method = "SCT",
                                         anchor.features = set.features)
  sets.integrated <- IntegrateData(anchorset = sets.anchors, normalization.method = "SCT")
  return(sets.integrated)
  
}

# for all

sets.integrated <- integrate_sets(sets.list)
saveRDS(object = sets.integrated, file = "output/integrated-all.RDS")

```








